---
title: "Project 1 Writeup"
author: "Mengqi Wang"
date: "Feb 8, 2016"
output: html_document
---

A chord diagram should illustrate intuitively the relationship of skills, i.e., the proportion of people who have a skill (e.g. SQL) also have anther skill (e.g. Python). Also it would be good for visualizing the relationship between skills and program of people.

### Visualization of program-skills relationship
To visualize this relationship, we need to selection features (columns corresponding to skillset questions in our case and program column), split each skill into one new column as bitmap (e.g. if 1 in SQL means familiarity for SQL and 0 means not). So the cleaning scripts as describe in previous sectors are used

First I further cleaned the data using following script:
```{r, echo=FALSE}
#Set your working directory
#setwd('Documents/cuw4701_edav/')
#Load CSV file into Data Frame
df = read.csv("edit_data.csv")
col.list = c("Matlab", "R", "Github", "Excel", "SQL", "RStudio", "ggplot2", "shell (terminal / command line)", "C/C++", "Python", "Stata", "LaTeX", "XML", "Web: html css js", "google drive (formerly docs)", "Sweave/knitr","dropbox", "SPSS", "regular expressions (grep)", "lattice" )

#Create colums initializing at 0
df[,col.list] = 0

for(i in col.list){
  #Need an If Statement because of R vs RStudio. 
  if(i == "R"){ 
    #Use Reg expressions "R,|R$" which looks for "R," and for "R$" which means there is nothing after R (line 87 caused this issue)
    fnd = "R,|R$"
    #try to find fnd within the vector, return Row # if True
    rows = grep(pattern = fnd, x = df$Experiences.with.tools)
  }else{
    #Same as above
    fnd = paste(i, sep = "")
    rows = grep(pattern = fnd, x = df$Experiences.with.tools, fixed = TRUE)
  }
  df[rows,i] = 1
}

separate_major <- function(survey){
  IDSE = which(survey$Program == "IDSE (master)")
  Other = which(survey$Program == "Other masters")
  DSC = which(survey$Program == "Data Science Certification")
  STATS = which(survey$Program == "Statistics (master)")
  PHD = which(survey$Program == "Ph.D.")
  
  a = list(IDSE, DSC, STATS, PHD, Other)
  return(a) 
}
```


Filter those skills that matters. In other words, if one skill has too few people, we filtered that out.
```{r, fig.high=12, fig.width=8, warning=FALSE}
#install.packages("circlize")
library(circlize)

# filter features
df_clean <- df[,c(12:31)]
colnames(df_clean)[8] = 'Shell'
colnames(df_clean)[14] = 'Web'
colnames(df_clean)[15] = 'Google Doc'
df_clean = df_clean[,-c(13, 16:20)]

#colnames(df_clean)[2] = "ProgrammingProficiency"
#df_clean[13, 1] = df_clean[1, 1]
majors = separate_major(df)
IDSE = colMeans(df_clean[majors[[1]],])
DSC = colMeans(df_clean[majors[[2]],])
STATS = colMeans(df_clean[majors[[3]],])
PHD = colMeans(df_clean[majors[[4]],])
Other = colMeans(df_clean[majors[[5]],])

df_cd = rbind(IDSE, DSC, STATS, PHD, Other)
df_bymajor = data.frame(from = rep(rownames(df_cd), times = ncol(df_cd)), to = rep(colnames(df_cd), each = nrow(df_cd)),
                value = as.vector(df_cd),
                stringsAsFactors = FALSE)
chordDiagram(df_bymajor)
```

So the chord diagram for program to skills is created. Each degree has a corresponding arc in the circle and each chord (the colorful thick lines inside the circle) connects a proportion of students in each program to their corresponding each skill.

### Visualization of skill-skill relationship
Then we further transformed the dataset for creating a new chord diagram showing the relationship skill-skill relationship.
```{r, fig.high=12, fig.width=8}
df_skillset = data.frame(matrix(rep(0,dim(df_clean)[2] * dim(df_clean)[2]), nrow=dim(df_clean)[2], ncol=dim(df_clean)[2]))
colnames(df_skillset) = colnames(df_clean)
rownames(df_skillset) = colnames(df_clean)
sk_list = colnames(df_clean)
for(i in 1:dim(df_clean)[1]) {
  for(j in 1:dim(df_clean)[2]) {
    for(k in 1:j) {
      if((df_clean[i, j] == 1 && df_clean[i, k] == 1) && (sk_list[j] != sk_list[k])) {
        df_skillset[j, k] = df_skillset[j, k] + 1
        # Set weight between same skill (Matlab, Matlab) as 0
      }
    }
  }
}

df_byskill = data.frame(from = rep(rownames(df_skillset), times = ncol(df_skillset)), to = rep(colnames(df_skillset), each = nrow(df_skillset)), value = as.vector(unlist(df_skillset)), stringsAsFactors = FALSE)
grid.col = NULL
grid.col[sk_list] = 1:length(sk_list)
chordDiagram(df_byskill, grid.col = grid.col)
```

