Yuhao Code
=============================
I created a visualization that shows the information of experience with tools across majors. There are two plots made. One compares stats with data science and the other compares the data science master with data science certificates. I chose these two majors because the majority of people are in these two majors. We can compare as many majors as we want if needed. The graph shows for each tool what is the proportion of people who know how to use the tool.The further away the lines from the center, the higher the proportion of people who can use that tool. I choose the top five tools identified in the random forest classifier. We can clearly see that the patterns for two majors are different. Also, we can use this kind of plot to see what are people from different good at. 

```{r}
setwd("/Users/vernon/Dropbox/courses/Columbia/STAT W4701 EDA and visualization/hw/hw1")
survey = read.csv("Survey+Response.csv",na.strings = "NA")
unique_majors = unique(survey$Program)

```
separate the experience_with_tools into bitmap

```{r}
df = read.csv("Survey+Response.csv")

col.list = c("Matlab", "R", "Github", "Excel", "SQL", "RStudio", "ggplot2", "shell (terminal / command line)", "C/C++", "Python", "Stata", "LaTeX", "XML", "Web: html css js", "google drive (formerly docs)", "Sweave/knitr","dropbox", "SPSS", "regular expressions (grep)", "lattice" )

#these are the top five tools defined in the Bob's random forest
important_skills = c("Python", "Github", "google drive (formerly docs)", "Matlab", "ggplot2")


#Create colums initializing at 0

df[,col.list] = 0
for(i in col.list){
  #Need an If Statement because of R vs RStudio. 
    if(i == "R"){ 
      #Use Reg expressions "R,|R$" which looks for "R," and for "R$" which means there is nothing after R (line 87 caused this issue)
      fnd = "R,|R$"
      #try to find fnd within the vector, return Row # if True
      rows = grep(pattern = fnd, x = df$Experiences.with.tools)
    }else{
      #Same as above
      fnd = paste(i, sep = "")
      rows = grep(pattern = fnd, x = df$Experiences.with.tools, fixed = TRUE)
      }
    df[rows,i] = 1
}
```
Now separate majors and do some calculation over the skills
```{r}

separate_major <- function(survey){
  IDSE = which(survey$Program == "IDSE (master)")
  Other = which(survey$Program == "Other masters")
  DSC = which(survey$Program == "Data Science Certification")
  STATS = which(survey$Program == "Statistics (master)")
  QMSS = which(survey$Program == "QMSS")
  DS = which(survey$Program == "Data Science")
  MS_DS = which(survey$Program == "Ms in ds")
  MS_QMSS = which(survey$Program == "QMSS (master)")
  PHD = which(survey$Program == "Ph.D.")
  APP_MATH = which(survey$Program == "Applied Math")
  PHD_BIO = which(survey$Program == "PhD Biomedical Informatics")
  MSDS = which(survey$Program == "MSDS")

  
  MSDS = c(MSDS, IDSE, DS, MS_DS)
  QMSS = c(QMSS, MS_QMSS)
  PHD = c(PHD, PHD_BIO)
  a = list(MSDS, QMSS, STATS, PHD, APP_MATH, Other, DSC)
  return(a) 
}
attempt = separate_major(df)


#col.list is the list of skills, skill is the output of function skill_distri, then skill[[1]] is the ratio of people who know the skill, skill[[2]] is number of people who know the skill, skill[[3]] is the number of people who do not know the skill
skill_distri <- function(major, col.list){
  total_num = dim(major)[1]
  skills_ratio = c()
  skill_count = c()
  differ = c()
  #among the people of the same major, the ratio of people who have a skill
  for (i in 1:length(col.list)){
    #print(col.list[i])
    num = sum(major[,col.list[i]])
    ratio = num/total_num
    diff = total_num - num
    skills_ratio = c(skills_ratio, ratio)
    skill_count = c(skill_count, num)
    differ = c(differ, diff)
  }
  return(list(skills_ratio, skill_count, differ))
}


MSDS = df[attempt[[1]],]
MSDS_Skills = skill_distri(MSDS, important_skills)
MSDS_result = data.frame(rbind(MSDS_Skills[[2]], MSDS_Skills[[3]]))
colnames(MSDS_result) = important_skills


STATS = df[attempt[[3]],]
STATS_Skills = skill_distri(STATS, important_skills)
STATS_result = data.frame(rbind(STATS_Skills[[2]], STATS_Skills[[3]]))
colnames(STATS_result) = important_skills

DSC = df[attempt[[7]],]
DSC_Skills = skill_distri(DSC, important_skills)
DSC_result = data.frame(rbind(DSC_Skills[[2]], DSC_Skills[[3]]))
colnames(DSC_result) = important_skills

```
radar plot
http://www.statisticstoproveanything.com/2013/11/spider-web-plots-in-r.html
```{r}

webplot = function(data, data.row = NULL, y.cols = NULL, main = NULL, add = F, 
    col = "red", lty = 1, scale = T) {
    if (!is.matrix(data) & !is.data.frame(data)) 
        stop("Requires matrix or data.frame")
    if (is.null(y.cols)) 
        y.cols = colnames(data)[sapply(data, is.numeric)]
    if (sum(!sapply(data[, y.cols], is.numeric)) > 0) {
        out = paste0("\"", colnames(data)[!sapply(data, is.numeric)], "\"", 
            collapse = ", ")
        stop(paste0("All y.cols must be numeric\n", out, "are not numeric"))
    }
    if (is.null(data.row)) #plot the data of the first row
        data.row = 1
    if (is.character(data.row)) 
        if (data.row %in% rownames(data)) {
            data.row = which(rownames(data) == data.row)
        } else {
            stop("Invalid value for data.row:\nMust be a valid rownames(data) or row-index value")
        }
    if (is.null(main)) 
        main = rownames(data)[data.row]
    if (scale == T) {
        #data = scale(data[, y.cols])
        #data = apply(data, 2, function(x) x/max(abs(x)))
        data = apply(data, 2, function(x) x/sum(abs(x))) #dataframe$col/sum(dataframe$col) = each element divided by the sum of corresponding column
    }
    data = as.data.frame(data)
    n.y = length(y.cols)
    min.rad = 360/n.y
    #polar.vals = (90 + seq(0, 360, length.out = n.y + 1)) * pi/180
    polar.vals = (90 + seq(0, 360, length.out = n.y + 1)) * pi/180

    # 
    if (add == F) {
        plot(0, xlim = c(-2.2, 2.2), ylim = c(-2.2, 2.2), type = "n", axes = F, 
            xlab = "", ylab = "", asp = 1)
        title(main)
        lapply(polar.vals, function(x) lines(c(0, 2 * cos(x)), c(0, 2 * sin(x))))
        lapply(1:n.y, function(x) text(2.15 * cos(polar.vals[x]), 2.15 * sin(polar.vals[x]), 
            y.cols[x], cex = 0.8))

        #lapply(seq(0.5, 2, 0.5), function(x) lines(x * cos(seq(0, 2 * pi, length.out = 100)), 
        #    x * sin(seq(0, 2 * pi, length.out = 100)), lwd = 0.5, lty =2, col = "gray60"))
        #lines(cos(seq(0, 2 * pi, length.out = 100)), sin(seq(0, 2 * pi, length.out = 100)), 
        #    lwd = 1.2, col = "gray50")
        
        lapply(seq(0.5, 2, 0.5), function(x) lines(x * cos(seq(0, 2 * pi, length.out = 100)), 
            x * sin(seq(0, 2 * pi, length.out = 100)), lwd = 0.5, lty = 2, col = "gray60"))
        lines(cos(seq(0, 2 * pi, length.out = 100)), sin(seq(0, 2 * pi, length.out = 100)), 
            lwd = 0.5, lty = 2, col = "gray50")
    }


    r = 1 + data[data.row, y.cols]
    xs = r * cos(polar.vals)
    ys = r * sin(polar.vals)
    xs = c(xs, xs[1])
    ys = c(ys, ys[1])

    lines(xs, ys, col = col, lwd = 2, lty = lty)

}

par(mar = c(1, 1, 2, 1))
webplot(MSDS_result, main = "Data Science VS Stats")
webplot(STATS_result, add = T, col = "blue", lty = 2)
par(new = T)
par(mar = c(0, 0, 0, 0))
plot(0, type = "n", axes = F)
legend("bottomright", lty = c(1, 2), lwd = 2, col = c("red", "blue"), c("MS.Data Science", 
    "Stats"), bty = "n")


par(mar = c(1, 1, 2, 1))
webplot(MSDS_result, main = "Data Science Master VS Certificate")
webplot(DSC_result, add = T, col = "blue", lty = 2)
par(new = T)
par(mar = c(0, 0, 0, 0))
plot(0, type = "n", axes = F)
legend("bottomright", lty = c(1, 2), lwd = 2, col = c("red", "blue"), c("MS.Data Science", 
    "DS certificate"), bty = "n")

```








